# Data transformation

```{r}
library(dplyr)
```

```{r}
rad_people <- read.csv("data/raw/People.csv")
colnames(rad_people)[1] <- "FIPS"
rad_jobs <- read.csv("data/raw/Jobs.csv")
colnames(rad_jobs)[1] <- "FIPS"
rad_income <- read.csv("data/raw/Income.csv")
colnames(rad_income)[1] <- "FIPS"
rad_CC <- read.csv("data/raw/County Classifications.csv")
colnames(rad_CC)[1] <- "FIPS"
```

```{r}
rad_people <- rad_people%>%select(FIPS,Ed1LessThanHSPct,Ed5CollegePlusPct,Ed4AssocDegreePct,Ed2HSDiplomaOnlyPct)
rad_CC <- rad_CC %>% select(FIPS,Hipov,HiAmenity,Low_Education_2015_update)
rad_income <- rad_income %>% select(-(County),-(State))
```

```{r}
rad22 <- full_join(rad_jobs, rad_people, by = "FIPS")%>%
              full_join(.,rad_income, by = "FIPS")%>%
                    full_join(.,rad_CC, by = "FIPS")%>%
                      distinct(FIPS, .keep_all = TRUE)
rad22$FIPS <- as.factor(rad22$FIPS)
```

We have 5 excel files, they are very easy to load into dataframes that are workable in R.\
We only kept 4 of them because the variables in veterans.csv does not pertain to the questions we are trying to answer.\
For some of the dataset, we only chose a subset of columns that are relevant (expand on this a little more, maybe use a few examples to explain why these columns are selected and why some others are discarded).\
The data types of the columns all make sense, so we don’t need to do much about it.\
We used factor for the state column to make future R analyses/graphing easier (optional) .\
Merged the (useful) dataframes together with the value of FIPs, since FIPs is a like an ID for each county, it is unique and we can identify each county with a FIPs value.\
Lastly, this new cleaned data is saved into a csv file in the folder data/cleaned, which will be used for all the analysis in the rest of this report.
